---
title: "MARKDOWN TEMPLATE"
author: "Nick Graetz"
date: "Sept 4, 2020"
output: pdf_document
---
  
```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# Load knitr package and settings
library(knitr)
library(data.table)
library(ggplot2)
library(gridExtra)
library(stargazer)
options(scipen=999)
knitr::opts_chunk$set(fig.width=8, fig.height=4, fig.align = "center") 
## Check out: https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf
```

# **Problem 1**

Let's do some math:

\begin{align}
Y&=\mu \\
&=\beta \\
&=2+2 
\end{align}

# **Problem 2**

Let's do some code:

```{r, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
x <- 10
d <- data.table(x=1:10,
                y=(1:10)^2)
print(x)
```

In my code I made a numeric vector `r x`.  

I can set the echo=TRUE argument in my R code chunk to print out the contents of the chunk in my final document.  

```{r, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
x <- 10
d <- data.table(a=1:10,
                b=(1:10)^2)
```

There are a bunch of packages to make pretty tables from data.frame type objects, kable() is the simplest

```{r, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
kable(d)
```

\pagebreak

# **Problem 3**

Let's do some code and make a plot:

```{r, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
ggplot(data=d) +
  geom_line(aes(x=a,
                y=b)) + 
  theme_bw()
```